Windowed least-squares model reduction for dynamical systems

### 2019-10-08 main.tex
latest

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a \methodNameLower\ (\methodAcronym) approach for model-reduction
	of dynamical systems. The proposed approach sequentially minimizes the
	time-continuous full-order-model residual within a low-dimensional space--time trial
	subspace over time windows. The approach comprises a generalization 
of existing model reduction approaches, as particular instances of
  the methodology recover the Galerkin,
	least-squares Petrov--Galerkin (LSPG), and space–time LSPG approaches. In
	addition, the approach
	addresses key deficiencies in existing model-reduction
	techniques, e.g., the dependence of the (space--time) LSPG approach on the
	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
	error bounds for both the Galerkin and LSPG approaches.  We consider two types of
	space--time trial
	subspaces within the proposed approach: one that reduces only the spatial dimension of the full-order model, and one that reduces both the spatial and temporal dimensions of the full-order model. For each type of trial
	subspace, we consider two different solution techniques: direct (i.e.,
	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
	Numerical experiments conducted using trial subspaces characterized by spatial
	dimension reduction demonstrate that the \methodAcronym\
	approach can yield more accurate solutions with lower space--time residuals than
	the Galerkin and LSPG approaches.

@@

Residual-minimization methods in model reduction compute the solution within a
low-dimensional trial subspace that minimizes the full-order-model residual.
Researchers have developed such residual-minimizing model-reduction methods
for both static systems (i.e., systems without time-dependence)
\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and dynamical
systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
In the latter category,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{time-discrete} full-order-model residual
(i.e., the residual arising after applying time discretization) at each time
instance on the time-discretization grid. This formulation is often referred
to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
formulation and examined its connections with the Galerkin approach. Critically,
this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
Petrov--Galerkin approach applied to the time-continuous
full-order-model residual, and (2) the LSPG and Galerkin approaches are equivalent
in the limit as the time step goes to zero (i.e., the Galerkin approach 
minimizes the time-instantaneous full-order-model residual). Numerous
numerical experiments have demonstrated that LSPG often yields more
accurate and stable solutions than Galerkin
~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common intuitive explanation for this improved performance is that, by
minimizing the full-order-model residual over a finite time window (rather
than time instantaneously), LSPG computes solutions that are more
accurate over a larger part of the trajectory as compared to Galerkin.

However, LSPG has several notable shortcomings. First, LSPG
exhibits a complex dependence on the time discretization. In
particular, changing
the time step ($\Delta t$) modifies both the
time window over which LSPG minimizes the residual as well as the
time-discretization error of the full-order model on which LSPG is based. 
As LSPG and Galerkin are equivalent in the
limit of $\Delta t \rightarrow 0$, the accuracy of LSPG approaches the (sometimes poor)
accuracy of Galerkin as the time step shrinks.
For too-large a time step the accuracy of LSPG also degrades. It is unclear if this is due to the time-discretization error associated with enlarging the time step, or rather if it is due to the size of the window the residual is being minimized over.
% but this also has the effect of
%reducing the time-dicretization error of the full-order model.
As a
consequence, LSPG often yields the smallest error for an intermediate value of the
time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}); there
is no known way to compute this optimal time step \textit{a priori}.
Second, as the LSPG approach performs sequential residual minimization in
time, its \textit{a posteriori} error bounds grow exponentially in time
\cite{carlberg_lspg_v_galerkin}, and it is not equipped with any notion of
optimality over the entire time domain of interest. As a result, LSPG is not
equipped with \textit{a priori} guarantees of accuracy or stability, even for
linear time-invariant systems~\cite{bui_thesis}.

Researchers have pursued the development of space--time
residual-minimization approaches~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
address the issues incurred by sequential residual minimization in time.
Existing space--time approaches differ from the classic LSPG and Galerkin approaches in (1) the definition of the space--time trial subspace and (2) the definition of the residual minimization problem. First, space--time approaches leverage a space--time trial 
subspace that associates with a space--time projection; classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage a space--time trial subspace that can be associated with spatial projection only. Second, space--time residual minimization approaches compute the entire space--time trajectory of the state (within the low-dimensional space--time trial subspace) that minimizes the full-order-model residual over the entire time domain; Galerkin and LSPG sequentially compute instances of the state that either minimize the full-order model instantaneously (Galerkin) or over a time-step instance (LSPG). 
The result of these differences is that space--time approaches yield an algebraic system of equations defined
over all space and time, whose solution comprises a \text{fixed} vector of
(space--time) generalized coordinates; on the other hand, spatial projection only approaches 
generally associates with a system of ODEs whose solution
comprises a \textit{time-dependent} vector of (spatial) generalized
coordinates.

@@

The objectives of this work are to overcome the shortcomings of existing
residual-minimizing methods, and to provide a unifying framework from which
existing methods can be assessed. In essence, the proposed  \textit{\methodNameLower\
(\methodAcronym)} method sequentially minimizes the FOM
residual over a sequence of arbitrarily defined
time windows. The method is characterized by three notable aspects.  First, the method
minimizes the \textit{time-continuous}  residual (i.e., that associated with
the full-order model ODE). By adopting a time-continuous viewpoint, the
formulation decouples the underlying temporal discretization scheme from the
residual-minimization problem, thus addressing a key deficiency of both LSPG
and space--time LSPG. Critically, time-continuous residual minimization also exposes
two different solution methods:  a \textit{discretize-then-optimize}
(i.e., direct) method, and an \textit{optimize-then-discretize} (i.e.,
indirect) method. Second, the method sequentially minimizes
the residual over arbitrarily defined \textit{time windows} rather than
sequentially minimizing the residual over time steps (as in LSPG)
or over the entire time domain (as in space--time residual-minimization
methods). This equips the method with additionally flexibility that enables it
to explore more fine-grained tradeoffs between computational cost and error.
Finally, \methodAcronym\ is formulated for two kinds of space--time trial
subspaces: one that associates with spatial projection only (as employed by traditional spatial-projection-based methods), and one that associates with space--time projection (as employed by space--time methods).
The above attributes allow the \methodAcronym\ approach to be viewed as a
generalization of existing model-reduction methods, as 
Galerkin, LSPG, and space--time LSPG correspond to specific
instances of the formulation.
Figure~\ref{fig:flowchart} depicts how the proposed \methodAcronym\ method
provides a unifying framework from which these existing approaches can be derived.

The \methodAcronym\ approach can be viewed as a hybrid space--time method and displays 
commonalities with several related efforts. 
First, Ref.~\cite{bui_thesis} briefly formulated a space--time least-squares ROM
and connected this formulation with optimal control; specifically it mentioned the
optimize-then-discretize vs.\ discretize-then-optimize approaches. This work
did not fully develop this approach, eschewing it 
for 
sequential residual minimization in time (i.e., LSPG). The present work thus
formally develops and extends several of the concepts put forth
in Ref.~\cite{bui_thesis}. Next, Ref.~\cite{constantine_strom} developed a
space--time residual minimization formulation for model interpolation.  The
present work distinguishes itself from Ref.~\cite{constantine_strom} in that
(1) this work also considers trial subspaces characterized by spatial projection only, and (2) we
minimize the time-continuous FOM residual over arbitrary time
windows. We note that,
similar to the current work, Ref.~\cite{constantine_strom} employs
minimization of the time-continuous FOM residual as a starting point; as such,
this work shares some thematic similarities with
Ref.~\cite{constantine_strom}.  Lastly, Ref.~\cite{choi_stlspg} develops a
space--time extension of the LSPG approach that minimizes the time-discrete
FOM residual over the entire time domain.  The present work distinguishes
itself from Ref.~\cite{choi_stlspg} in that (1) this work minimizes the
time-continuous FOM residual, (2) this work minimizes this residual over
arbitrary time windows, and (3) this work also considers trial subspaces
associated with spatial projection only.

@@

The paper proceeds as follows: Section~\ref{sec:math}
outlines the mathematical setting for the full-order model, along with the Galerkin, LSPG, and
space--time LSPG approaches. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
approach. Section~\ref{sec:numerical_techniques}
outlines numerical techniques for solving \methodAcronymROMs, including both direct and
indirect methods. Section~\ref{sec:analysis} provides
equivalence conditions and error analysis for \methodAcronymROMs.
Section~\ref{sec:numerical_experiments} presents numerical experiments.
Section~\ref{sec:conclude} provides conclusions and perspectives.
We denote vector-valued functions with italicized bold symbols (e.g., $\boldsymbol
x$), vectors with standard bold symbols (e.g., $\mathbf{x}$), 
matrices with capital bold symbols (e.g., $\mathbf{X} \equiv \begin{bmatrix}
\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
function with respect to time with the $\dot\ $ operator.
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical formulation}\label{sec:math}
	We begin by providing the formulation for the full-order model,
	followed by a description of standard model-reduction methods
	classified according to the type of trial subspace they employ.
	\subsection{Full-order model}\label{sec;FOM}
We consider the full-order model to be a dynamical system expressed as a
	system of ordinary differential equations (ODEs)
\begin{equation}\label{eq:FOM}
 \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
	\stateFOMIC,\qquad t \in [0,T],
\end{equation}
where  $\stateFOM: [0,T]
	\rightarrow  \RR{\fomdim}$ with $\stateFOM: \timeDummy \mapsto
	\stateFOM(\timeDummy) $ and $\stateFOMDot \equiv {d \stateFOM}/{d\tau}$
	denotes the state implicitly defined as the solution to initial value
	problem~\eqref{eq:FOM}, 
$T \in \RRplus$ denotes the final time, 
 $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDummy) \mapsto
	\velocity(\stateyDiscrete,\timeDummy)$ denotes the velocity, which is possibly
	nonlinear in its first argument. For subsequent exposition, we introduce
	$\timeSpace$ to denote the set of (sufficiently smooth) real-valued functions acting on the time
	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
	state can be expressed equivalently as $\stateFOM \in \RR{\fomdim} \otimes
	\timeSpace $.  We refer to the initial value problem defined in
	Eq.~\eqref{eq:FOM} as the ``full-order model" (FOM) ODE. We note that
	although the problem of interest described in the introduction corresponds
	to a parameterized dynamical system, we suppress dependence of the FOM ODE
	\eqref{eq:FOM} on such parameters for notational convenience, as this work
	focuses on devising a model-reduction approach applicable to a specific parameter instance.

@@

\subsection{Reduced-order models}
Projection-based ROMs generate approximate solutions to the FOM
	ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial
	subspace. Two types of space--time trial subspaces are commonly used for
	this purpose:\footnote{For both spatial and space--time ROMs of dynamical systems, all trial subspaces are, strictly speaking, space--time subspaces.} 
\begin{enumerate} 
	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
		model (\spatialAcronym)}. These trial subspaces are characterized by a \textit{spatial projection} and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
%	trial subspaces are commonly used
%		in model-reduction methods based on spatial projection followed by time
%		integration (e.g., the Galerkin and LSPG approaches).
	\item \textit{Subspaces that reduce both the spatial and temporal dimensions of the full-order
		model (\spaceTimeAcronym)}.
These trial subspaces are characterized by a \textit{space--time} projection and are employed in space--time 
model reduction approaches (e.g., space--time Galerkin~\cite{benner_st}, space--time LSPG~\cite{choi_stlspg}). 
% These trial subspaces are commonly employed in
%		model-reduction methods based on space--time projection (e.g., space--time
%		LSPG).
\end{enumerate}
 We now describe these two types of space--time trial subspaces and their
	application to the Galerkin, LSPG, and space--time LSPG approaches.

@@

\subsubsection{Galerkin approach}
The Galerkin approach reduces the number of equations in
Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the 
residual to the spatial trial subspace in the (semi-)inner product induced by the 
positive (semi-)definite $\fomdim\times\fomdim$ matrix $\stweightingMatArg{}\equiv \stweightingMatOneTArg{ }
\stweightingMatOneArg{ }$ (commonly set to
$\stweightingMatArg{}=\mathbf{I}$), i.e.,
\begin{equation}\label{eq:g_truncation2}
\genstateGalerkinDotArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace
\genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
\end{equation}
where $\massArgnt{ } \equiv \basisspace^T \stweightingMatArg{ } \basisspace$
denotes the $K\times K$ positive definite mass matrix.
As demonstrated in Ref.~\cite{carlberg_lspg_v_galerkin}, the Galerkin approach
can be viewed alternatively as a residual-minimization
method, as the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent
to
\begin{equation}\label{eq:GalOptimal}
\genstateGalerkinDotArg{}{t} = \underset{ \genstateyDiscrete \in \RR{\romdim}
	}{\text{arg\,min}} \| \basisspace \hat{\stateyDiscrete} -
	\velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t)
	\|_{\stweightingMatArg{}}^2, \qquad \genstate(0) = \genstateIC,
	\qquad t \in [0,T],
%\hat{\mathbf{v}}, %\qquad \genstateGalerkin(0) = \basisspace^T(\stateFOMIC -\stateIntercept), \qquad t \in [0,T].
\end{equation}
where $\|\mathbf{x}\|_{\stweightingMatArg{}}\equiv
\sqrt{\mathbf{x}^T\stweightingMatArg{}\mathbf{x}}$.
Thus, the computed velocity $\genstateGalerkinDotArg{}{t}$ minimizes the
FOM ODE residual evaluated at the state $\basisspace
\genstateGalerkin(t) + \stateIntercept$ and time instance $t$ over the spatial trial
subspace $\trialspace$.
%To perform dimension reduction, 
%To perform dimension reduction, the governing equations are projected by a ``test" basis,
%\begin{equation}\label{eq:g_truncation}
%\testbasis^T \basisspace \frac{d \genstate}{dt} = \testbasis^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
%The most common choice is to set the test space and trial space to be equivalent, which results in the Galerkin reduced-order model,
%\begin{equation}\label{eq:g_truncation}
%\frac{d \genstate}{dt} = \basisspace^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
\subsubsection{The least-squares Petrov--Galerkin approach}
Despite its time-instantaneous residual-minimization optimality property 
\eqref{eq:GalOptimal}, the Galerkin approach can yield inaccurate solutions,
particularly if the velocity is not self-adjoint or
is nonlinear. 
Least-squares Petrov--Galerkin (LSPG)
~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
was developed
as an alternative method that exhibits several advantages over the Galerkin approach.
Rather than minimize the (time-continuous) FOM ODE residual at a time instance (as in Galerkin), LSPG minimizes
the (time-discrete) FOM O$\Delta$E residual 
(i.e., the residual arising after applying time discretization to the FOM ODE)
over a time step.
We now describe the LSPG approach in the case of linear multistep methods;
Ref.~\cite{carlberg_lspg_v_galerkin} also presents LSPG for Runge--Kutta
schemes.

>>>>

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a \methodNameLower\ (\methodAcronym) approach for model-reduction
	of dynamical systems. The proposed approach sequentially minimizes the
	time-continuous full-order-model residual within a low-dimensional space--time trial
	subspace over time windows. The approach comprises a generalization 
of existing model reduction approaches, as particular instances of
  the methodology recover Galerkin,
	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection. In
	addition, the approach
	addresses key deficiencies in existing model-reduction
	techniques, e.g., the dependence of (space--time) LSPG projection on the
	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
	error bounds for both Galerkin and LSPG projection.  We consider two types of
	space--time trial
	subspaces within the proposed approach: one that reduces only the spatial dimension of the full-order model, and one that reduces both the spatial and temporal dimensions of the full-order model. For each type of trial
	subspace, we consider two different solution techniques: direct (i.e.,
	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
	Numerical experiments conducted using trial subspaces characterized by spatial
	dimension reduction demonstrate that the \methodAcronym\
	approach can yield more accurate solutions with lower space--time residuals than
	Galerkin and LSPG projection.

@@

Residual-minimization methods in model reduction compute the solution within a
low-dimensional trial subspace that minimizes the full-order-model residual.
Researchers have developed such residual-minimizing model-reduction methods
for both static systems (i.e., systems without time-dependence)
\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and dynamical
systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
In the latter category,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{time-discrete} full-order-model residual
(i.e., the residual arising after applying time discretization) at each time
instance on the time-discretization grid. This formulation is often referred
to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
formulation and examined its connections with Galerkin projection. Critically,
this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
Petrov--Galerkin projection applied to the time-continuous
full-order-model residual, and (2) the LSPG and Galerkin projection are equivalent
in the limit as the time step goes to zero (i.e., Galerkin projection 
minimizes the time-instantaneous full-order-model residual). Numerous
numerical experiments have demonstrated that LSPG often yields more
accurate and stable solutions than Galerkin
~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common intuitive explanation for this improved performance is that, by
minimizing the full-order-model residual over a finite time window (rather
than time instantaneously), LSPG computes solutions that are more
accurate over a larger part of the trajectory as compared to Galerkin.

However, LSPG has several notable shortcomings. First, LSPG
exhibits a complex dependence on the time discretization. In
particular, changing
the time step ($\Delta t$) modifies both the
time window over which LSPG minimizes the residual as well as the
time-discretization error of the full-order model on which LSPG is based. 
As LSPG and Galerkin projection are equivalent in the
limit of $\Delta t \rightarrow 0$, the accuracy of LSPG approaches the (sometimes poor)
accuracy of Galerkin as the time step shrinks.
For too-large a time step the accuracy of LSPG also degrades. It is unclear if this is due to the time-discretization error associated with enlarging the time step, or rather if it is due to the size of the window the residual is being minimized over.
% but this also has the effect of
%reducing the time-dicretization error of the full-order model.
As a
consequence, LSPG often yields the smallest error for an intermediate value of the
time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}); there
is no known way to compute this optimal time step \textit{a priori}.
Second, as the LSPG approach performs sequential residual minimization in
time, its \textit{a posteriori} error bounds grow exponentially in time
\cite{carlberg_lspg_v_galerkin}, and it is not equipped with any notion of
optimality over the entire time domain of interest. As a result, LSPG is not
equipped with \textit{a priori} guarantees of accuracy or stability, even for
linear time-invariant systems~\cite{bui_thesis}.

Researchers have pursued the development of space--time
residual-minimization approaches~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC} to
address the issues incurred by sequential residual minimization in time.
Existing space--time approaches differ from the classic LSPG and Galerkin approaches in (1) the definition of the space--time trial subspace and (2) the definition of the residual minimization problem. First, space--time approaches leverage a space--time trial 
subspace that associates with a space--time projection operator; classic ``spatial" model reduction approaches such as LSPG and Galerkin leverage a space--time trial subspace that can be associated with a spatial projection operator only. Second, space--time residual minimization approaches compute the entire space--time trajectory of the state (within the low-dimensional space--time trial subspace) that minimizes the full-order-model residual over the entire time domain; Galerkin and LSPG sequentially compute instances of the state that either minimize the full-order model instantaneously (Galerkin) or over a time-step instance (LSPG). 
The result of these differences is that space--time approaches yield an algebraic system of equations defined
over all space and time, whose solution comprises a \text{fixed} vector of
(space--time) generalized coordinates; on the other hand, spatial projection only approaches 
generally associates with a system of ODEs whose solution
comprises a \textit{time-dependent} vector of (spatial) generalized
coordinates.

@@

The objectives of this work are to overcome the shortcomings of existing
residual-minimizing methods, and to provide a unifying framework from which
existing methods can be assessed. In essence, the proposed \textit{\methodNameLower\
(\methodAcronym)} \approachKwd\ sequentially minimizes the FOM
residual over a sequence of arbitrarily defined
time windows. The method is characterized by three notable aspects.  First, the method
minimizes the \textit{time-continuous}  residual (i.e., that associated with
the full-order model ODE). By adopting a time-continuous viewpoint, the
formulation decouples the underlying temporal discretization scheme from the
residual-minimization problem, thus addressing a key deficiency of both LSPG
and space--time LSPG. Critically, time-continuous residual minimization also exposes
two different solution methods:  a \textit{discretize-then-optimize}
(i.e., direct) method, and an \textit{optimize-then-discretize} (i.e.,
indirect) method. Second, the method sequentially minimizes
the residual over arbitrarily defined \textit{time windows} rather than
sequentially minimizing the residual over time steps (as in LSPG)
or over the entire time domain (as in space--time residual-minimization
methods). This equips the method with additionally flexibility that enables it
to explore more fine-grained tradeoffs between computational cost and error.
Finally, \methodAcronym\ is formulated for two kinds of space--time trial
subspaces: one that associates with spatial projection only (as employed by traditional spatial-projection-based methods), and one that associates with space--time projection (as employed by space--time methods).
The above attributes allow the \methodAcronym\ approach to be viewed as a
generalization of existing model-reduction methods, as 
Galerkin, LSPG, and space--time LSPG projection correspond to specific
instances of the formulation.
Figure~\ref{fig:flowchart} depicts how the proposed \methodAcronym\ method
provides a unifying framework from which these existing approaches can be derived.

The \methodAcronym\ approach can be viewed as a hybrid space--time method and displays 
commonalities with several related efforts. 
First, Ref.~\cite{bui_thesis} briefly formulated a space--time least-squares ROM
and connected this formulation with optimal control; specifically it mentioned the
optimize-then-discretize vs.\ discretize-then-optimize approaches. This work
did not fully develop this approach, eschewing it 
for 
sequential residual minimization in time (i.e., LSPG). The present work thus
formally develops and extends several of the concepts put forth
in Ref.~\cite{bui_thesis}. Next, Ref.~\cite{constantine_strom} developed a
space--time residual minimization formulation for model interpolation.  The
present work distinguishes itself from Ref.~\cite{constantine_strom} in that
(1) this work also considers trial subspaces characterized by spatial projection only, and (2) we
minimize the time-continuous FOM residual over arbitrary time
windows. We note that,
similar to the current work, Ref.~\cite{constantine_strom} employs
minimization of the time-continuous FOM residual as a starting point; as such,
this work shares some thematic similarities with
Ref.~\cite{constantine_strom}.  Lastly, Ref.~\cite{choi_stlspg} develops a
space--time extension of LSPG projection that minimizes the time-discrete
FOM residual over the entire time domain.  The present work distinguishes
itself from Ref.~\cite{choi_stlspg} in that (1) this work minimizes the
time-continuous FOM residual, (2) this work minimizes this residual over
arbitrary time windows, and (3) this work also considers trial subspaces
associated with spatial projection only.

@@

The paper proceeds as follows: Section~\ref{sec:math}
outlines the mathematical setting for the full-order model, along with Galerkin, LSPG, and
space--time LSPG projection. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
\approachKwd. Section~\ref{sec:numerical_techniques}
outlines numerical techniques for solving \methodAcronymROMs, including both direct and
indirect methods. Section~\ref{sec:analysis} provides
equivalence conditions and error analysis for \methodAcronymROMs.
Section~\ref{sec:numerical_experiments} presents numerical experiments.
Section~\ref{sec:conclude} provides conclusions and perspectives.
We denote vector-valued functions with italicized bold symbols (e.g., $\boldsymbol
x$), vectors with standard bold symbols (e.g., $\mathbf{x}$), 
matrices with capital bold symbols (e.g., $\mathbf{X} \equiv \begin{bmatrix}
\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
function with respect to time with the $\dot\ $ operator.
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical formulation}\label{sec:math}
	We begin by providing the formulation for the full-order model,
	followed by a description of standard model-reduction methods
	classified according to the type of trial subspace they employ.
	\subsection{Full-order model}\label{sec;FOM}
We consider the full-order model to be a dynamical system expressed as a
	system of ordinary differential equations (ODEs)
\begin{equation}\label{eq:FOM}
 \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
	\stateFOMIC,\qquad t \in [0,T],
\end{equation}
where  $\stateFOM: [0,T]
	\rightarrow  \RR{\fomdim}$ with $\stateFOM: \timeDummy \mapsto
	\stateFOM(\timeDummy) $ and $\stateFOMDot \equiv {d \stateFOM}/{d\tau}$
	denotes the state implicitly defined as the solution to initial value
	problem~\eqref{eq:FOM}, 
$T \in \RRplus$ denotes the final time, 
 $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDummy) \mapsto
	\velocity(\stateyDiscrete,\timeDummy)$ denotes the velocity, which is possibly
	nonlinear in its first argument. For subsequent exposition, we introduce
	$\timeSpace$ to denote the set of (sufficiently smooth) real-valued functions acting on the time
	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
	state can be expressed equivalently as $\stateFOM \in \RR{\fomdim} \otimes
	\timeSpace $.  We refer to the initial value problem defined in
	Eq.~\eqref{eq:FOM} as the ``full-order model" (FOM) ODE. We note that
	although the problem of interest described in the introduction corresponds
	to a parameterized dynamical system, we suppress dependence of the FOM ODE
	\eqref{eq:FOM} on such parameters for notational convenience, as this work
	focuses on devising a model-reduction approach applicable to a specific parameter instance.

@@

\subsection{Reduced-order models}
Projection-based ROMs generate approximate solutions to the FOM
	ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial
	subspace. Two types of space--time trial subspaces are commonly used for
	this purpose:\footnote{For both spatial and space--time ROMs of dynamical systems, all trial subspaces are, strictly speaking, space--time subspaces.} 
\begin{enumerate} 
	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
		model (\spatialAcronym)}. These trial subspaces are characterized by a \textit{spatial projection operator} and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
%	trial subspaces are commonly used
%		in model-reduction methods based on spatial projection followed by time
%		integration (e.g., the Galerkin and LSPG approaches).
	\item \textit{Subspaces that reduce both the spatial and temporal dimensions of the full-order
		model (\spaceTimeAcronym)}.
These trial subspaces are characterized by a \textit{space--time projection operator} and are employed in space--time 
model reduction approaches (e.g., space--time Galerkin~\cite{benner_st}, space--time LSPG~\cite{choi_stlspg}). 
% These trial subspaces are commonly employed in
%		model-reduction methods based on space--time projection (e.g., space--time
%		LSPG).
\end{enumerate}
 We now describe these two types of space--time trial subspaces and their
	application to the Galerkin, LSPG, and space--time LSPG approaches.

@@

\subsubsection{Galerkin projection}
The Galerkin approach reduces the number of equations in
Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the 
residual to the spatial trial subspace in the (semi-)inner product induced by the 
positive (semi-)definite $\fomdim\times\fomdim$ matrix $\stweightingMatArg{}\equiv \stweightingMatOneTArg{ }
\stweightingMatOneArg{ }$ (commonly set to
$\stweightingMatArg{}=\mathbf{I}$), i.e.,
\begin{equation}\label{eq:g_truncation2}
\genstateGalerkinDotArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace
\genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
\end{equation}
where $\massArgnt{ } \equiv \basisspace^T \stweightingMatArg{ } \basisspace$
denotes the $K\times K$ positive definite mass matrix.
As demonstrated in Ref.~\cite{carlberg_lspg_v_galerkin}, the Galerkin approach
can be viewed alternatively as a residual-minimization
method, as the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent
to
\begin{equation}\label{eq:GalOptimal}
\genstateGalerkinDotArg{}{t} = \underset{ \genstateyDiscrete \in \RR{\romdim}
	}{\text{arg\,min}} \| \basisspace \hat{\stateyDiscrete} -
	\velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t)
	\|_{\stweightingMatArg{}}^2, \qquad \genstate(0) = \genstateIC,
	\qquad t \in [0,T],
%\hat{\mathbf{v}}, %\qquad \genstateGalerkin(0) = \basisspace^T(\stateFOMIC -\stateIntercept), \qquad t \in [0,T].
\end{equation}
where $\|\mathbf{x}\|_{\stweightingMatArg{}}\equiv
\sqrt{\mathbf{x}^T\stweightingMatArg{}\mathbf{x}}$.
Thus, the computed velocity $\genstateGalerkinDotArg{}{t}$ minimizes the
FOM ODE residual evaluated at the state $\basisspace
\genstateGalerkin(t) + \stateIntercept$ and time instance $t$ over the spatial trial
subspace $\trialspace$.
%To perform dimension reduction, 
%To perform dimension reduction, the governing equations are projected by a ``test" basis,
%\begin{equation}\label{eq:g_truncation}
%\testbasis^T \basisspace \frac{d \genstate}{dt} = \testbasis^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
%The most common choice is to set the test space and trial space to be equivalent, which results in the Galerkin reduced-order model,
%\begin{equation}\label{eq:g_truncation}
%\frac{d \genstate}{dt} = \basisspace^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
\subsubsection{Least-squares Petrov--Galerkin projection}
Despite its time-instantaneous residual-minimization optimality property 
\eqref{eq:GalOptimal}, the Galerkin approach can yield inaccurate solutions,
particularly if the velocity is not self-adjoint or
is nonlinear. 
Least-squares Petrov--Galerkin (LSPG)
~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
was developed
as an alternative method that exhibits several advantages over the Galerkin approach.
Rather than minimize the (time-continuous) FOM ODE residual at a time instance (as in Galerkin), LSPG minimizes
the (time-discrete) FOM O$\Delta$E residual 
(i.e., the residual arising after applying time discretization to the FOM ODE)
over a time step.
We now describe the LSPG approach in the case of linear multistep methods;
Ref.~\cite{carlberg_lspg_v_galerkin} also presents LSPG for Runge--Kutta
schemes.
